Interval Probability for Fuzzy Quantum Theories

### 2016-07-24 prob.tex
prob.tex \subsection{Quantum Interval-valued Probability Measures} cleaning up...

Although there is no easy Radon-Nikodym theorem, an interval-valued
probability measure can still be understood by its relation with real-valued
probability measures, like we constructed example~\ref{ex3} from
example~\ref{ex1}. In general, a \emph{core} of an interval-valued
probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is
a set of real-valued probability measures~$\pmeas:\events\rightarrow[0,1]$
such that $\pmeas\left(E\right)\in\bar{\mu}\left(E\right)$ for all
$E\in\events$~\cite{GilboaSchmeidler1994,Marinacci1999,Teran2014}.

@@

%%%%%
\subsection{Quantum Probability Measures}

@@

Notice that if the last condition is replaced by the equal sign, these
conditions will be exactly the same as the conditions in lemma~\ref{lem:quantumProbabilityMeasure}
which is equivalent to the Born rule as in corollary~\ref{cor:Gleason's}.
Therefore, it is reasonable to believe these conditions should be
equivalent to a interval-valued Born rule.

When we pick $\mathscr{I}=\left\{ \left[x,x\right]\middle|x\in\left[0,1\right]\right\} $,
for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$,
we can define a corresponding quantum \emph{$\mathscr{I}$}-interval-valued
probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
by $\bar{\mu}\left(P\right)=\left[\mu\left(P\right),\mu\left(P\right)\right]$.
In this sense, every usual quantum probability measure can be considered
as a quantum interval-valued probability measure. There are more interesting
examples as we now show.

@@

Although $\iota:\left[0,1\right]\rightarrow\mathscr{I}$ maps every
quantum probability measure to a quantum interval-valued probability
measure, there is a quantum \emph{$\mathscr{I}$}-interval-valued
probability measure which does not come from a quantum probability
measure.

@@

Notice that in the above two examples, quantum interval-valued probability
measures are all come from quantum probability measures because the
numbers in $\left[0,1\right]$ can be mapped to the chosen intervals~$\mathscr{I}$
naturally. This is not always the case. For example, if $\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
there is no natural way to map from $\frac{1}{2}$, as both $\frac{1}{2}\in\left[0,\frac{1}{2}\right]=\unlikely$
and $\frac{1}{2}\in\left[\frac{1}{2},1\right]=\likely$. However,
we still can find a quantum interval-valued probability measure corresponding
to a quantum probability measure as in the following example.

\begin{example}[One-qubit quantum interval-valued probability measure]
Given a two dimensional Hilbert space, a vector $\ket{\psi}$ can
be normalized and written in $\left(\begin{array}{c}
\cos\theta\\
\rme^{\rmi\gamma}\sin\theta
\end{array}\right)$, where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. Consider
a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
defined by 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0}) & = & \imposs\textrm{ ,}\\
\bar{\mu}(\mathbb{1}) & = & \necess\textrm{ ,}\\
\bar{\mu}(\proj{\psi}) & = & \begin{cases}
\necess & \textrm{, if }\theta=0\textrm{ ;}\\
\likely & \textrm{, if }0<\theta<\frac{\pi}{4}\textrm{ ;}\\
\likely & \textrm{, if }\theta=\frac{\pi}{4}\textrm{ and }0\le\gamma<\pi\textrm{ ;}\\
\unlikely & \textrm{, if }\theta=\frac{\pi}{4}\textrm{ and }\pi\le\gamma<2\pi\textrm{ ;}\\
\unlikely & \textrm{, if }\frac{\pi}{4}<\theta<\frac{\pi}{2}\textrm{ ;}\\
\imposs & \textrm{, if }\theta=\frac{\pi}{2}\textrm{ .}
\end{cases}
\end{eqnarray*}
Let $\ket{\psi^{\perp}}$ be the state perpendicular to $\ket{\psi}$.
Then, the vector representation of $\ket{\psi^{\perp}}$ can be simplified
as 
\[
\left(\begin{array}{c}
-\rme^{-\rmi\gamma}\sin\theta\\
\cos\theta
\end{array}\right)=-\rme^{-\rmi\gamma}\left(\begin{array}{c}
\cos\left(\frac{\pi}{2}-\theta\right)\\
\rme^{\rmi\left(\pi+\gamma\right)}\sin\left(\frac{\pi}{2}-\theta\right)
\end{array}\right)
\]
so we have 
\[
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi}\right)=\bar{\mu}\left(\proj{\psi^{\perp}}\right)=1\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(\proj{\psi}\right)\textrm{ .}
\]
Hence, $\bar{\mu}$ is a quantum $\mathscr{I}$-interval-valued probability
measure. Notice that for all projection~$P$, $\mu_{\ket{0}}^{B}(P)\in\bar{\mu}(P)$
implies that $\bar{\mu}(P)$ is one of quantum $\mathscr{I}$-interval-valued
probability measures corresponding to $\mu_{\ket{0}}^{B}(P)$. However,
$\mu_{\ket{0}}^{B}(P)$ can correspond to infinity many quantum $\mathscr{I}$-interval-valued
probability measures by twisting the definition involving $\gamma$.\qed\end{example}

@@

\amr{
We can use DQC if we have some kind of topology (distances). The idea
will be that we want to prepare state PSI but because of errors etc we
prepare a close state. Well the next closest state will be the next
state in our discrete grid. I am sure that a state that's very close
to PSI can involve some wrapping around.
}

\yutsung{About continuity, one of the confusion comes from the subtle
correspondence between the measurement processes and observables.
Consider the following observables and their measurement processes:
\begin{itemize}
\item For any $x\in\mathbb{R}$, consider the observable operator~$\mathbf{O}_{x}=\proj{1}+x\proj{-1}$.
Except $x=1$, given the system in the state~$\ket{\phi}$, we can
denote the measurement process for $\mathbf{O}_{x}$ by ${\cal O}_{x}$:
\begin{enumerate}
\item With probability~$\mu_{\phi}\left(\proj{1}\right)$, return the measurement
result~$1$ with the post-measurement state~$\ket{1}$;
\item With probability~$\mu_{\phi}\left(\proj{0}\right)$, return the measurement
result~$0$ with the post-measurement state~$\ket{0}$;
\item With probability~$\mu_{\phi}\left(\proj{-1}\right)$, return the
measurement result~$x$ with the post-measurement state~$\ket{-1}$.
\end{enumerate}
\item The observable operator~$\mathbf{J}_{z}^{2}=\proj{1}+\proj{-1}$,
and its measurement process~${\cal J}_{z}^{2}$:
\begin{enumerate}
\item With probability~$\mu_{\phi}\left(\proj{1}+\proj{-1}\right)$, return
the measurement result~$1$ with the post-measurement state~$\ket{1}\ip{1}{\phi}+\ket{-1}\ip{-1}{\phi}$;
\item With probability~$\mu_{\phi}\left(\proj{0}\right)$, return the measurement
result~$0$ with the post-measurement state~$\ket{0}$.
\end{enumerate}
\end{itemize}
As an operator, it is clear that $\lim_{x\rightarrow1}\mathbf{O}_{x}=\mathbf{J}_{z}^{2}$,
and we should have $\lim_{x\rightarrow1}{\cal O}_{x}={\cal O}_{1}$
for the measurement process. However, it is not obvious why ${\cal O}_{1}={\cal J}_{z}^{2}$.
Although they have the same probability for the measurement results,
but their post-measurement states are different. This might be a defect
of representing a measurement process as an observable. As $\mathbf{O}_{1}=\mathbf{J}_{z}^{2}$,
there is no observable corresponding to the measurement process~${\cal O}_{1}$.}

\amr{the rest needs cleaning up and perhaps does not even belong in
  this section}

Although it seems that we need an infinite long table to specify the
quantum probability measure~$\mu$, our $\mu$ is actually
given by a simple formula~$\melement{0}{P}$. In general, Born discovered
each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\mu^B_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $P\in\qevents$ as follows~\cite{Born1983,Mermin2007}:
\begin{equation}
\mu^B_{\psi}(P)=\frac{\melement{\psi}{P}}{\ip{\psi}{\psi}}\label{eq:Born}
\end{equation}
The Born rule satisfies the following properties:
\begin{itemize}
\item It can be extend to mixed states. Given a mixed state represented
by a density matrix $\rho=\sum_{j=1}^{N}q_{j}\frac{\proj{\psi_{j}}}{\ip{\psi_{j}}{\psi_{j}}}$,
where $\sum_{j=1}^{N}q_{j}=1$, i.e., $\Tr\left(\rho\right)=1$, then
the Born rule can be extended to $\rho$ by 
\begin{eqnarray}
\mu^B_{\rho}\left(P\right) & = & \Tr\left(\rho P\right)=\sum_{j=1}^{N}q_{j}\mu^B_{\Psi_{j}}\left(P\right)\textrm{ .}\label{BornRule.mixed}
\end{eqnarray}
Notice that $\left(\left\{ 1,\ldots,N\right\} ,2^{\left\{ 1,\ldots,N\right\} },\pmeas\left(J\right)=\sum_{j\in J}q_{j}\right)$
is a classical probability space. Therefore, when we discretize the
Hilbert space later, we may need to discretize this probability space
as well.
\item $\mu^B_{\rho}$ is a probability measure for all mixed state~$\rho$.
\item $\ip{\psi}{\phi}=0\Leftrightarrow\mu^B_{\psi}\left(\proj{\phi}\right)=0$.
\item $\mu^B_{\psi}\left(P\right)=\mu^B_{\mathbf{U}\ket{\psi}}\left(\mathbf{U}P\mathbf{U}^{\dagger}\right)\textrm{ ,}$where
$\mathbf{U}$ is any unitary map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbb{1}$. 
\end{itemize}

Naturally, we may ask: is every probability measure induced from a
state by the Born rule? The answer is yes by Gleason's theorem when
the dimension~$\ge3$~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.
Furthermore, a simple corollary of Gleason's theorem can show the
Born rule is the unique function satisfying conditions 1. to 3.
\begin{cor}
The Born rule is the unique function satisfying conditions 1. to 3.
\end{cor}
\begin{proof}
Assume there is another function $\mu^{\prime B}$ such that $\mu^{\prime B}_{\rho}$
is a quantum probability measure for all mixed state~$\rho$. We
are going to prove $\mu^{\prime B}=\mu^B$.

Fix a pure normalized state $\phi$, $\mu^{\prime B}_{\phi}$ is a quantum
probability measure by condition 2. By Gleason's theorem, there is
a mixed state ~$\rho'$, such that $\mu^{\prime B}_{\phi}\left(P\right)=\Tr\left(\rho'P\right)=\sum_{j=1}^{N}q_{j}\mu^B_{\psi_{j}}\left(P\right)$
for all event $P$.

Consider the event $P'=\mathbb{1}-\proj{\phi}$, we have 
\begin{eqnarray*}
0 & \overset{\textrm{Condition 3}}{=} & \mu^B_{\phi}\left(P'\right)\\
 & = & \sum_{j=1}^{N}q_{j}\mu^B_{\psi_{j}}\left(P'\right)
\end{eqnarray*}
Because $q_{j}>0$, we have $\mu^B_{\psi_{j}}\left(P\right)=0$,
i.e., $\psi_{j}$ is orthogonal to a co-dimension-$1$ subspace $P'$.
However, the only subspace orthogonal to $P'$ is span by $\ket{\phi}$.
Hence, $\mu^{\prime B}_{\phi}=\mu^B_{\phi}$.
\end{proof}

% If there may be more than one probability measure, we will discuss
% whether we will keep using the Born rule (\ref{eq:Born}) or there
% is another formula $\mu^B$ such that $\mu^B_{\psi}$
% is a probability measure for all $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $.

% Then, Gleason's theorem states given a probability measure $\mu$
% there is a mixed state $\ket{\psi}$ such that $\mu=\mu_{\psi}$,
% i.e., the Born rule is surjective\footnote{If we extend the domain of $\mu_{\psi}$ including the mixed states.}.
% For any other formula~$\mu^B$, we can ask whether $\mu^B$
% is surjective as well.

% Obviously, we don't want arbitrarily assign a state~$\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
% with a probability measure $\mu^B_{\psi}$, for example, assigning
% every state to the same probability measure. We want $\mu^B_{\psi}$
% satisfying the following properties with some physical meaning:
% \begin{itemize}
% \item $\ip{\psi}{\phi}=0\Leftrightarrow\mu^B_{\psi}\left(\proj{\phi}\right)=\tilde{0}$,
% where $\tilde{0}$ is $0$ for $\left[0,1\right]$ and $\tilde{0}$
% is impossible for $\mathscr{L}_{2}=\left\{ \imposs,\poss\right\} $. 
% \item $\mu^B_{\psi}\left(\proj{\phi}\right)=\mu^B_{\mathbf{U}\ket{\psi}}\left(\mathbf{U}\proj{\phi}\mathbf{U}^{\dagger}\right)\textrm{ ,}$where
% $\ket{\psi},\ket{\phi}$ are states and $\mathbf{U}$ is any unitary
% map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbb{1}$. 
% \end{itemize}
% And the results can be summarized in the following table:\\
% \center{ %
% \begin{tabular}{>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}}
% \hline 
% State space $\Hilb$  & Probability values  & Is there a $\mu^B$ satisfying the given conditions?  & Is the $\mu^B$ surjective?\tabularnewline
% \hline 
% \hline 
% $\C^{d}$ for $d\ge3$  & $\left[0,1\right]$  & Yes  & Yes\tabularnewline
% \hline 
% $\C^{d}$  & $\mathscr{L}_{2}$  & Yes  & No\tabularnewline
% \hline 
% $\ffzd{p^{2}}$for $d\ge3$ except $d=p=3$  & $\left[0,1\right]$  & No  & \tabularnewline
% \hline 
% $\ffzd{p^{2}}$  & $\mathscr{L}_{2}$  & Yes & No\tabularnewline
% \hline 
% \end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{All Continuous or All Discrete}

Before we turn to the main part of the paper, we quickly dismiss the
possibility of having one but not the other of the discrete
variations. Specifically, it is impossible to maintain the Hilbert
space and have a finite set-valued probability measure and it is also
impossible to have a vector space constructed over a finite field with
a real-valued probability measure.

%%%
\subsection{Hilbert Space with Finite Set-Valued Probability Measure}

However, there is a $\mathscr{L}_{2}$-valued probability measure
\[
\hat{\mu}_{1}\left(P\right)=\begin{cases}
\imposs & \textrm{, if }P=\proj{+};\\
\bar{\mu}(P) & \textrm{, otherwise.}
\end{cases}
\]
such that $\hat{\mu}_{1}\ne\bar{\mu}_{\psi}$ for all mixed state
$\ket{\psi}$.

%%%
\subsection{Discrete Vector Space with Real-Valued Probability Measure}

>>>>

Although there is no easy Radon-Nikodym theorem, an interval-valued
probability measure can still be understood by its relation with real-valued
probability measures, like we constructed example~\ref{ex3} from
example~\ref{ex1}. In general, this idea can be generalized to the
following definition of core.

\begin{definition}[Core of Interval-valued Probability Measure~\cite{GilboaSchmeidler1994,Marinacci1999,Teran2014}]\label{def:CoreInterval-valuedProbability}
A \emph{core} of an interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
is the set 
\[
\mathrm{core}\left(\bar{\mu}\right)=\left\{ \pmeas:\events\rightarrow[0,1]\textrm{ probability measure}\middle|\pmeas\left(E\right)\in\bar{\mu}\left(E\right)\forall E\in\events\right\} \textrm{ .}
\]
\qed\end{definition}

@@

%%%%%
\subsection{Quantum Real-valued Probability Measures}

@@

Similar to the classical interval-valued probability measure, a quantum
interval-valued probability measure can be constructed from a real-valued
one.

@@

When a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
constructed from a real-valued one~$\mu:\events\rightarrow\left[0,1\right]$,
we should say $\mu$ is in the core of $\bar{\mu}$ according to the
following definition of core.

\begin{definition}[Core of Quantum Interval-valued Probability
Measure]\label{def:CoreQuantumInterval-valuedProbability} A \emph{core}
of a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
is the set 
\[
\mathrm{core}\left(\bar{\mu}\right)=\left\{ \mu:\events\rightarrow[0,1]\textrm{ quantum probability measure}\middle|\mu\left(P\right)\in\bar{\mu}\left(P\right)\forall P\in\events\right\} \textrm{ .}
\]
\qed\end{definition}

Because the classical interval-valued probability measure has no easy
Radon-Nikodym theorem, the quantum one should not have a easy Gleason's
theorem. Nevertheless, the core of $\bar{\mu}$ provides an instructive
way to understand $\bar{\mu}$, and the core may exist even without
a natural $\iota:\left[0,1\right]\rightarrow\mathscr{I}$ as in the
example~\ref{ex:One-qubit-quantum-interval} in appendix. However,
in contrast to the classical case, we have the following theorem.

\begin{thm}\label{def:EmptyCoreQuantumInterval-valuedProbability}
There is a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that $\mathrm{core}\left(\bar{\mu}\right)=\emptyset$.\qed\end{thm}

This theorem can be proved by the following example, and a more sophisticated
example is example~\ref{ex:three-dimensional-4-value} in appendix.

@@

\appendix

\section{Examples for Quantum Interval-valued Probability Measures}

The following example verify the core of a quantum interval-valued
probability measure is nonempty even if there is no natural correspondence.
For example, if $\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
there is no natural way to map from $\frac{1}{2}$, as both $\frac{1}{2}\in\left[0,\frac{1}{2}\right]=\unlikely$
and $\frac{1}{2}\in\left[\frac{1}{2},1\right]=\likely$. However,
we still can find a quantum interval-valued probability measure corresponding
to a quantum probability measure as in the following example.

\begin{example}[One-qubit quantum interval-valued probability measure]\label{ex:One-qubit-quantum-interval}
Given a two dimensional Hilbert space, a vector $\ket{\psi}$ can
be normalized and written in $\left(\begin{array}{c}
\cos\theta\\
\rme^{\rmi\gamma}\sin\theta
\end{array}\right)$, where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. Consider
a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
defined by 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0}) & = & \imposs\textrm{ ,}\\
\bar{\mu}(\mathbb{1}) & = & \necess\textrm{ ,}\\
\bar{\mu}(\proj{\psi}) & = & \begin{cases}
\necess & \textrm{, if }\theta=0\textrm{ ;}\\
\likely & \textrm{, if }0<\theta<\frac{\pi}{4}\textrm{ ;}\\
\likely & \textrm{, if }\theta=\frac{\pi}{4}\textrm{ and }0\le\gamma<\pi\textrm{ ;}\\
\unlikely & \textrm{, if }\theta=\frac{\pi}{4}\textrm{ and }\pi\le\gamma<2\pi\textrm{ ;}\\
\unlikely & \textrm{, if }\frac{\pi}{4}<\theta<\frac{\pi}{2}\textrm{ ;}\\
\imposs & \textrm{, if }\theta=\frac{\pi}{2}\textrm{ .}
\end{cases}
\end{eqnarray*}
Let $\ket{\psi^{\perp}}$ be the state perpendicular to $\ket{\psi}$.
Then, the vector representation of $\ket{\psi^{\perp}}$ can be simplified
as 
\[
\left(\begin{array}{c}
-\rme^{-\rmi\gamma}\sin\theta\\
\cos\theta
\end{array}\right)=-\rme^{-\rmi\gamma}\left(\begin{array}{c}
\cos\left(\frac{\pi}{2}-\theta\right)\\
\rme^{\rmi\left(\pi+\gamma\right)}\sin\left(\frac{\pi}{2}-\theta\right)
\end{array}\right)
\]
so we have 
\[
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi}\right)=\bar{\mu}\left(\proj{\psi^{\perp}}\right)=1\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(\proj{\psi}\right)\textrm{ .}
\]
Hence, $\bar{\mu}$ is a quantum $\mathscr{I}$-interval-valued probability
measure. Notice that for all projection~$P$, $\mu_{\ket{0}}^{B}(P)\in\bar{\mu}(P)$
implies that $\bar{\mu}(P)$ is one of quantum $\mathscr{I}$-interval-valued
probability measures corresponding to $\mu_{\ket{0}}^{B}(P)$. However,
$\mu_{\ket{0}}^{B}(P)$ can correspond to infinity many quantum $\mathscr{I}$-interval-valued
probability measures by twisting the definition involving $\gamma$.\qed\end{example}

@@
